---
title: "4. Estimate vegetation properties from optical images"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{4. Estimate vegetation properties from optical images}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial is based on the estimation of vegetation properties from 
Sentinel-2 data and `prosail` hybrid inversion. 
It provides the full framework step by step, including Sentinel-2 data download, 
image preprocessing and masking, hybrid model training, and the application of 
the model on an image. 


# Identifying and downloading Sentinel-2 images

The Sentinel-2 acquisition corresponds to the Barrax area in Spain. 
Barrax is a well-studied area for remote sensing applications on crop monitoring. 
A polygon corresponding to the study area can be downloaded from a repository. 

```{r identify_S2_product}
# define bounding box for area of interest
aoi_bbox <- sf::st_bbox(obj = c('xmin' = -2.25, 'ymin' = 38.94, 
                                'xmax' = -1.73, 'ymax' = 39.22))
sf::st_crs(x = aoi_bbox) <- 4326

```

[Sentinel-hub](https://apps.sentinel-hub.com/eo-browser/?zoom=11&lat=39.09836&lng=-2.01462&themeId=DEFAULT-THEME&visualizationUrl=https%3A%2F%2Fservices.sentinel-hub.com%2Fogc%2Fwms%2Fbd86bcc0-f318-402b-a145-015f85b9427e&datasetId=S2L2A&fromTime=2021-05-13T00%3A00%3A00.000Z&toTime=2021-05-13T23%3A59%3A59.999Z&layerId=1_TRUE_COLOR) 
is then used in order to identify an acquisition with minimum cloud cover.

Once a date of acquisition is identified, the S2 product is then downloaded. 
The R package [`preprocS2`](https://jbferet.gitlab.io/preprocs2/) is used for this purpose. 
`preprocS2` will download S2 data from a STAC catalog based on spatial and temporal information.
The default STAC collection is `sentinel-2-l2a` from the Planetary computer catalog.

Optionally, `preprocS2` can download the geometry of acquisition corresponding to the S2 acquisition.
Access to geometry of acquisition is provided by the package [`CDSE`](https://zivankaraman.github.io/CDSE/).

```{r download_S2_product}
# library
library(preprocS2)
library(sf)

# 1- define input & output directories
output_dir_s2 <- './S2_images'
dir.create(path = output_dir, showWarnings = F, recursive = T)
datetime <- as.Date('2021-05-13')

# get tiling grid kml from https://sentiwiki.copernicus.eu/web/s2-products
path_S2tilinggrid <- 'Sentinel-2_tiling_grid.kml'

# 3- get S2 acquisition and geometry of acquisition for study area
# https://shapps.dataspace.copernicus.eu/dashboard/#/account/settings
id <- "sh-XXXXXXXX-XXXX-XXXX-XXXX-XXXXXXXXXXXX"
secret <- "XXXXXXXXXXXXXXXXXXXXXXXXXXXX"
authentication <- list('id' = id, 'pwd' = secret)

list_files <- get_s2_raster(bbox = aoi_bbox, datetime = datetime, output_dir = output_dir_s2, 
              path_S2tilinggrid = path_S2tilinggrid, siteName = 'Barrax',
              overwrite = F, geomAcq = T, authentication = authentication)
# Sentinel-2 L2A reflectance
Refl_L2A <- list_files$Refl_L2A
# Sentinel-2 binary mask identifying vegetation, discarding clouds & shadows
vegetation_mask <- list_files$vegetation_mask
# Sentinel-2 mask from provider ( = SCL from ESA products)
SCL <- list_files$provider_mask
```

# Training and application of hybrid inversion model 

```{r Hybrid_Inversion}
library(prosail)
output_dir_BP <- './PROSAIL'
dir.create(path = output_dir_BP, showWarnings = FALSE,recursive = TRUE)
# get sensor response for Sentinel-2
SensorName <- 'Sentinel_2'
SRF <- GetRadiometry(SensorName)

# define parameters to estimate
Parms2Estimate <- c('lai', 'CHL', 'EWT', 'LMA', 'fCover', 'fAPAR', 'albedo')

# define spectral bands required to train SVR model for each variable
S2BandSelect <- Bands2Select <- list()
for (parm in Parms2Estimate){
  S2BandSelect[[parm]] <- c('B3','B4','B5','B6','B7','B8','B11','B12')
  Bands2Select[[parm]] <- match(S2BandSelect[[parm]],SRF$Spectral_Bands)
}
ImgBandNames <- SRF$Spectral_Bands

# get S2 geometry: read metadata file from S2 image
angles <- c('saa', 'sza', 'vaa', 'vza')
angles_path <- file.path(output_dir_s2, 'geomAcq_S2', 
                         paste(angles, '_', datetime, '.tiff', sep = ''))
names(angles_path) <- angles
geom <- MeanAngle <- list()
for (angle in angles){
  geom[[angle]] <- terra::crop(x = terra::rast(angles_path[angle]), 
                               y = as(aoi_bbox, "Spatial"))
  MeanAngle[[angle]] <- mean(terra::values(geom[[angle]]))
}
MeanAngle$psi <- abs(MeanAngle$vaa - MeanAngle$saa)
if (MeanAngle$psi>180)
  MeanAngle$psi <- 360-MeanAngle$psi

# define ranges for zenith angles of acquisition
GeomAcq <- list('min' = data.frame('tto' = MeanAngle$vza,
                                   'tts' = MeanAngle$sza, 
                                   'psi' = MeanAngle$psi), 
                'max' = data.frame('tto' = MeanAngle$vza, 
                                   'tts' = MeanAngle$sza,
                                   'psi' = MeanAngle$psi))

# train regression model
# set method = 'svmRadial' or method = 'svmLinear' to use SVR implemented in the 
# R package caret instead of liquidSVM
modelSVR <- train_prosail_inversion(Parms2Estimate = Parms2Estimate,
                                    atbd = TRUE, GeomAcq = GeomAcq, 
                                    SRF = SRF, 
                                    Bands2Select = Bands2Select, 
                                    Path_Results = PROSAIL_ResPath, 
                                    method = 'liquidSVM')

# apply regression model on Sentinel-2 raster data
BPvars <- Apply_prosail_inversion(raster_path = Refl_path, 
                                  HybridModel = modelSVR, 
                                  PathOut = PROSAIL_ResPath,
                                  SelectedBands = S2BandSelect, 
                                  bandname = ImgBandNames, 
                                  MaskRaster = Cloud_File, 
                                  bigRaster = T, 
                                  MultiplyingFactor = 10000)
```


Assessed biophysical properties computed with `Apply_prosail_inversion` are then 
saved as individual raster files. 
For each biophysical variable, the mean value and standard deviation defined by 
the input variable `nbModels` from `modelSVR` (default = 20 models) is saved.
