---
title: "6. PROSAIL inversion with iterative optimization"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
  html_vignette:
    number_sections: true
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{6. PROSAIL inversion with iterative optimization}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```


PROSAIL inversion is ill-posed. 
Therefore, standard iterative optimization algorithms consisting in minimizing a 
cost function will converge towards local minimum dependent on initial 
conditions, unless global optimization algorithms are applied, or prior 
information is used. 

Inversion strategies based on iterative optimization with or without prior 
information are implemented in `prosail`. 


## Iterative optimization with no prior information

The main iterative optimization algorithm is based on the minimization of a 
multivariable function with nonlinear constraints, and uses the function 
`fmincon` included in the package `pracma`.
The default criterion for the cost function corresponds to the root mean square 
error (RMSE) between a reference surface reflectance and surface reflectance 
simulated with PROSAIL. 
Iterative optimization is applied to simulated surface reflectance at native 
spectral sampling (1 nm), and to Sentinel-2 surface reflectance as follows:

```{r iterative_no_prior}
# define initial value, lower and upper bounds for inversion
init <- data.frame('chl' = 40, 'car' = 10, 'ewt' = 0.01, 'lma' = 0.01, 
                   'lai' = 3, 'lidf_a' = 50, 'n_struct' = 1.5)
lb <- data.frame('chl' = 5, 'car' = 1, 'ewt' = 0.002, 'lma' = 0, 
                 'lai' = 0.5, 'lidf_a' = 30, 'n_struct' = 1)
ub <- data.frame('chl' = 80, 'car' = 20, 'ewt' = 0.03, 'lma' = 0.03, 
                 'lai' = 6, 'lidf_a' = 80, 'n_struct' = 3)

# define parameters set for inversion
parm_set <- data.frame('tts' = 40, 'tto' = 0, 'psi' = 60, 'soil_brightness' = 1,
                       'ant' = 0, 'brown' = 0, 'hotspot' = 0.1)

# define soil reflectance (set as prior knowledge) and spec_soil variable
rsoil <- parm_set$soil_brightness*spec_soil_ossl$soil_01
spec_soil <- data.frame('lambda' = spec_soil$lambda, 
                        'refl' = rsoil)

# simulate canopy surface reflectance with 1 nm sampling
truth <- data.frame('chl' = 60, 'car' = 8, 'ewt' = 0.015, 'lma' = 0.005,
                    'lai' = 3,  'lidf_a' = 60, 'n_struct' = 1.8)
refl_1nm <- prosail(n_struct = truth$n_struct, chl = truth$chl, car = truth$car,
                    ant = parm_set$ant, brown = parm_set$brown, 
                    ewt = truth$ewt, lma = truth$lma, type_lidf = 2, 
                    lai = truth$lai, hotspot = parm_set$hotspot, 
                    lidf_a = init$lidf_a, rsoil = rsoil, tts = parm_set$tts, 
                    tto = parm_set$tto, psi = parm_set$psi)
surf_refl_1nm <- get_surf_refl(rdot = refl_1nm$rdot, rsot = refl_1nm$rsot,
                               tts = parm_set$tts, 
                               spec_atm_sensor = spec_atm)

# invert PROSAIL on surface reflectance with 1 nm spectral sampling
est_1nm <- invert_prosail(refl_mes = surf_refl_1nm$surf_refl, 
                          initialization = init,
                          lower_bound = lb, upper_bound = ub, type_lidf = 2, 
                          spec_prospect_sensor = spec_prospect_full_range,
                          spec_atm_sensor = spec_atm, 
                          spec_soil_sensor = spec_soil, 
                          parm_set = parm_set)

# convert spectral input based on Sentinel-2 SRF
srf_s2 <- get_srf_sensor(sensor_name = 'Sentinel_2A')
spec_prospect <- spec_prospect_full_range
lambda <- spec_prospect_full_range$lambda
surf_refl_S2 <- apply_sensor_characteristics(wvl = lambda, 
                                             srf = srf_s2, 
                                             refl = surf_refl_1nm$surf_refl)
spec_prospect_s2 <- apply_sensor_characteristics(wvl = lambda, srf = srf_s2, 
                                                 refl = spec_prospect)
spec_soil_s2 <- apply_sensor_characteristics(wvl = lambda, srf = srf_s2, 
                                             refl = spec_soil)
spec_atm_s2 <- apply_sensor_characteristics(wvl = lambda, srf = srf_s2, 
                                            refl = spec_atm)

# invert PROSAIL on Sentinel-2 surface reflectance
est_s2 <- invert_prosail(refl_mes = surf_refl_S2, initialization = init,
                         lower_bound = lb, upper_bound = ub,
                         spec_prospect_sensor = spec_prospect_s2,
                         spec_atm_sensor = spec_atm_s2, 
                         spec_soil_sensor = spec_soil_s2,
                         type_lidf = 2, parm_set = parm_set)

```

## Iterative optimization with prior information

Prior information can be added to the cost function in order to regularize the 
ill-posed inverse problem ([Combal et al., 2003](https://doi.org/10.1016/S0034-4257(02)00035-4)).
Multiple possibilities exist to account for prior information: 

- the definition of `initialization` variable according to initial assumptions

- the addition of a term corresponding to the probability density function of 
part or all of the parameters to be estimated in the cost function. 

The code blow illustrates the introduction of prior information on LIDF, with the 
definition of a probability density function corresponding to the average leaf 
angle `lidf_a`. 
A weighting factor associated to the relative importance of this prior 
information with respect to the main criterion to minimize (RMSE between 
reference surface reflectance and simulated surface reflectance) can also be 
adjusted.


```{r iterative_prior}
# define prior information on average leaf angle lidf_a
prior_info <- list('mean' = data.frame('lidf_a' = 60), 
                   'sd' = data.frame('lidf_a' = 10), 
                   'weight_prior' = 0.01)

# invert PROSAIL on surface reflectance with 1 nm spectral sampling and 
# probability density function for lidf_a
est_1nm_p <- invert_prosail(refl_mes = surf_refl_1nm$surf_refl, 
                            initialization = init,
                            lower_bound = lb, upper_bound = ub,
                            spec_prospect_sensor = spec_prospect_full_range,
                            spec_atm_sensor = spec_atm, 
                            spec_soil_sensor = spec_soil,
                            type_lidf = 2, parm_set = parm_set,
                            prior_info =  prior_info)

# invert PROSAIL on Sentinel-2 surface reflectance and probability density 
# function for lidf_a
est_s2_p <- invert_prosail(refl_mes = surf_refl_S2, initialization = init,
                           lower_bound = lb, upper_bound = ub,
                           spec_prospect_sensor = spec_prospect_s2,
                           spec_atm_sensor = spec_atm_s2, 
                           spec_soil_sensor = spec_soil_s2,
                           type_lidf = 2, parm_set = parm_set,
                           prior_info =  prior_info)

```
