---
title: "4. Estimate vegetation properties from optical images"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{4. Estimate vegetation properties from optical images}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This tutorial is based on the estimation of vegetation properties from 
Sentinel-2 data and `prosail` hybrid inversion. 
It provides the full framework step by step, including Sentinel-2 data download, 
image preprocessing and masking, hybrid model training, and the application of 
the model on an image. 


# Identifying and downloading Sentinel-2 images

The Sentinel-2 acquisition corresponds to the Barrax area in Spain. 
Barrax is a well-studied area for remote sensing applications on crop monitoring. 
A polygon corresponding to the study area can be downloaded from a repository. 

```{r identify_S2_product}
# library
library(zip)
# define data directory
Path_Data <- '01_DATA'
dir.create(Path_Data,showWarnings = F, recursive = T)
# name zip file including plots located on the tile
destzip <- file.path(Path_Data,'Study_area_Barrax.zip')
# url for the zip file
url <- 'https://gitlab.com/jbferet/myshareddata/-/raw/master/preprocS2_Example/Study_area_Barrax.zip'
download.file(url = url, destfile = destzip)
destunz <- file.path(Path_Data,'Study_area_Barrax')
unzip(zipfile = destzip,exdir = destunz)
```

[Sentinel-hub](https://apps.sentinel-hub.com/eo-browser/?zoom=11&lat=39.09836&lng=-2.01462&themeId=DEFAULT-THEME&visualizationUrl=https%3A%2F%2Fservices.sentinel-hub.com%2Fogc%2Fwms%2Fbd86bcc0-f318-402b-a145-015f85b9427e&datasetId=S2L2A&fromTime=2021-05-13T00%3A00%3A00.000Z&toTime=2021-05-13T23%3A59%3A59.999Z&layerId=1_TRUE_COLOR) 
is then used in order to identify an acquisition with minimum cloud cover.

Once a date of acquisition is identified, the S2 product is then downloaded. 
The R package [`preprocS2`](https://jbferet.gitlab.io/preprocs2/) is used 
in order to identify the S2 product available for the date of acquisition over 
the study area, and to download the corresponding Sentinel-2 SAFE archive from 
the google Cloud Sentinel-2 bucket.
`preprocS2` is based on the R package [`sen2r`](https://sen2r.ranghetti.info/). 

__WARNING#1: [sen2r](https://github.com/ranghetti/sen2r/issues/464) is no longer maintained.__
__WARNING#2: Users are required to download S2 images from the Google Cloud Sentinel-2 bucket.__ 


```{r download_S2_product}
library(preprocS2)
# define date of S2 acquisition
dateAcq <- '2022-05-28'
# define path for study area
path_vector <- '01_DATA/Study_area_Barrax/Study_area_Barrax.shp'
# define output directory where SAFE zipfile is stored
DirWrite <- '01_DATA/S2_Images'
dir.create(DirWrite,showWarnings = F, recursive = T)
# download Sentinel-2 image corresponding to study area and date of acquisition 
Path_S2 <- get_S2_L2A_Image(l2a_path = DirWrite, 
                            spatial_extent = path_vector, 
                            dateAcq = dateAcq)
```

# Preprocess S2 image

Once the SAFE S2 image is downloaded, the R package 
[`preprocS2`](https://jbferet.gitlab.io/preprocS2/index.html) is used to stack 
individual bands, crop to required extent defined by the vector file, produce a 
binary cloud mask and save in the desired raster format.

The function `extract_from_S2_L2A` performs cropping and resampling to 10m if 
needed, and provides a stars object as output, along with location for metadata 
and image full name.

```{r preprocess_S2}
##____________________________________________________________________##
##        Define where data is stored and where to write results      ##
##--------------------------------------------------------------------##
# Result directory
result_path <- '03_RESULTS'
dir.create(path = result_path,showWarnings = FALSE,recursive = TRUE)

##____________________________________________________________________##
##                  Extract, resample & stack data                    ##
##--------------------------------------------------------------------##
# define resolution
resolution <- 10
# define source of data
S2source <- 'SAFE'
S2obj <- preprocS2::extract_from_S2_L2A(Path_dir_S2 = Path_S2,
                                        path_vector = path_vector,
                                        S2source = S2source,
                                        resolution = resolution)
# update shapefile if needed (reprojection)
path_vector <- S2obj$path_vector
```


# Write reflectance, cloud mask and metadata

Once the image is read, it can be written as stacked raster file, and stored 
with the binary cloud mask and the metadata file. 

```{r write_raster}
# create specific result directory corresponding to granule name
results_site_path <- file.path(result_path,basename(S2obj$S2_Bands$GRANULE))
dir.create(path = results_site_path,showWarnings = FALSE,recursive = TRUE)
##____________________________________________________________________##
##                        Write CLOUD MASK                            ##
##--------------------------------------------------------------------##
# directory for cloud mask
Cloud_path <- file.path(results_site_path,'CloudMask')
dir.create(path = Cloud_path,showWarnings = FALSE,recursive = TRUE)
# Filename for cloud mask
cloudmasks <- preprocS2::save_cloud_s2(S2_stars = S2obj$S2_Stack,
                                       Cloud_path = Cloud_path,
                                       footprint = path_vector,
                                       S2source = S2source, SaveRaw = T)
##____________________________________________________________________##
##                        Write REFLECTANCE                           ##
##--------------------------------------------------------------------##
# directory for Reflectance
Refl_dir <- file.path(results_site_path,'Reflectance')
dir.create(path = Refl_dir,showWarnings = FALSE,recursive = TRUE)
# filename for Reflectance
Refl_path <- file.path(Refl_dir,paste(basename(S2obj$S2_Bands$GRANULE),'_Refl',sep = ''))

# Save Reflectance file as ENVI image with BIL interleaves
tile_S2 <- preprocS2::get_tile(S2obj$S2_Bands$GRANULE)
dateAcq_S2 <- preprocS2::get_date(S2obj$S2_Bands$GRANULE)
preprocS2::save_reflectance_s2(S2_stars = S2obj$S2_Stack, 
                               Refl_path = Refl_path,
                               S2Sat = NULL, 
                               tile_S2 = tile_S2, 
                               dateAcq_S2 = dateAcq_S2,
                               Format = 'ENVI', 
                               datatype = 'Int16', 
                               MTD = S2obj$S2_Bands$metadata, 
                               MTD_MSI = S2obj$S2_Bands$metadata_MSI)
```


# masking non vegetated pixels with NDVI thresholding

The [`spinR`](https://gitlab.com/jbferet/spinr) package includes a set of 
functions for the computation of spectral indices.
We will compute NDVI in order to update the vegetation mask and discard
non-vegetated pixels. 


```{r Spectral_indices}
########################################################################
##                      COMPUTE SPECTRAL INDEX                        ##
########################################################################
library(spinR)
library(raster)
library(stars)
# Read raster
Refl <- brick(Refl_path)
# get raster band name and clean format. Expecting band name and wavelength to be documented in image
HDR_Refl <- read_ENVI_header(get_HDR_name(Refl_path))
SensorBands <- HDR_Refl$wavelength
# compute a set of spectral indices defined by IndexList from S2 data
IndexList <- c('NDVI')
# ReflFactor = 10000 when reflectance is coded as INT16
Indices <- spinR::compute_S2SI_Raster(Refl = Refl, SensorBands = SensorBands,
                                      Sel_Indices = IndexList,
                                      ReflFactor = 10000, StackOut=F)

# create directory for Spectral indices
SI_path <- file.path(results_site_path,'SpectralIndices')
dir.create(path = SI_path,showWarnings = FALSE,recursive = TRUE)
# Save spectral indices
for (SpIndx in names(Indices$SpectralIndices)){
  Index_Path <- file.path(SI_path,paste(basename(S2obj$S2_Bands$GRANULE),'_',SpIndx,sep = ''))
  stars::write_stars(st_as_stars(Indices$SpectralIndices[[SpIndx]]), dsn=Index_Path, driver =  "ENVI",type='Float32')
  # write band name in HDR
  HDR <- read_ENVI_header(get_HDR_name(Index_Path))
  HDR$`band names` <- SpIndx
  write_ENVI_header(HDR = HDR,HDRpath = get_HDR_name(Index_Path))
}

# Update Cloud mask based on radiometric filtering
# eliminate pixels with NDVI < NDVI_Thresh because not enough vegetation
NDVI_Thresh <- 0.5
Elim <- which(values(Indices$SpectralIndices[['NDVI']])<NDVI_Thresh)
CloudInit <- stars::read_stars(cloudmasks$BinaryMask)
CloudInit$CloudMask_Binary[Elim] <- 0
# save updated cloud mask
Cloud_File <- file.path(Cloud_path,'CloudMask_Binary_Update')
stars::write_stars(CloudInit, dsn=Cloud_File,driver = "ENVI",type='Byte')
```

# Training and application of hybrid inversion model 

```{r Hybrid_Inversion}
<<<<<<< HEAD

########################################################################
##      COMPUTE BIOPHYSICAL VARIABLES BASED ON PROSAIL INVERSION      ##
########################################################################
# get S2 geometry
# read metadata file from S2 image
xmlfile <- file.path(dirname(Refl_path),'MTD_TL.xml')
S2Geom <- get_S2geometry(MTD_TL_xml = xmlfile)

# Train PROSAIL inversion
minval <- data.frame('CHL'=10,'CAR'=0,'EWT' = 0.005,'ANT' = 0,'LMA' = 0.005,'N' = 1.0,'psoil' = 0.0, 'BROWN'=0.0,
                     'LIDFa' = 30, 'lai' = 0.5,'q'=0.1,'tto' = 0,'tts' = min(S2Geom$SZA), 'psi' = 5)
maxval <- data.frame('CHL'=90,'CAR'=20,'EWT' = 0.04,'ANT' = 3,'LMA' = 0.04,'N' = 2.0, 'psoil' = 1.0, 'BROWN'=0.5,
                     'LIDFa' = 70, 'lai' = 7,'q'=0.25,'tto' = 7,'tts' = max(S2Geom$SZA), 'psi' = 355)

=======
library(prosail)
>>>>>>> dev
# get sensor response for Sentinel-2
SensorName <- 'Sentinel_2'
SRF <- GetRadiometry(SensorName)

# define parameters to estimate
Parms2Estimate <- c('lai', 'CHL', 'EWT', 'LMA', 'fCover', 'fAPAR', 'albedo')

# define spectral bands required to train SVR model for each variable
S2BandSelect <- Bands2Select <- list()
for (parm in Parms2Estimate){
  S2BandSelect[[parm]] <- c('B3','B4','B5','B6','B7','B8','B11','B12')
  Bands2Select[[parm]] <- match(S2BandSelect[[parm]],SRF$Spectral_Bands)
}
# ImgBandNames <- strsplit(HDR_Refl$`band names`,split = ',')[[1]]
ImgBandNames <- c('B2','B3','B4','B5','B6','B7','B8','B11','B12')

# define output directory where LUTs will be saved
PROSAIL_ResPath <- file.path(results_site_path,'HybridInversion')
dir.create(path = PROSAIL_ResPath, showWarnings = FALSE,recursive = TRUE)

# get S2 geometry: read metadata file from S2 image
xmlfile <- file.path(dirname(Refl_path),'MTD_TL.xml')
S2Geom <- get_S2geometry(MTD_TL_xml = xmlfile)

# define ranges for zenith angles of acquisition
GeomAcq <- list()
GeomAcq$min <- data.frame('tto' = min(S2Geom$VZA), 'tts' = min(S2Geom$SZA), 
                          'psi' = 0)
GeomAcq$max <- data.frame('tto' = max(S2Geom$VZA), 'tts' = max(S2Geom$SZA), 
                          'psi' = 360)

# train regression model
modelSVR <- train_prosail_inversion(Parms2Estimate = Parms2Estimate,
                                    atbd = TRUE, GeomAcq = GeomAcq, 
                                    SRF = SRF, 
                                    Bands2Select = Bands2Select, 
                                    Path_Results = PROSAIL_ResPath)

# apply regression model on Sentinel-2 raster data
Apply_prosail_inversion(raster_path = Refl_path, 
                        HybridModel = modelSVR, 
                        PathOut = PROSAIL_ResPath,
                        SelectedBands = S2BandSelect, 
                        bandname = ImgBandNames, 
                        MaskRaster = Cloud_File, 
                        MultiplyingFactor = 10000)
```


Assessed biophysical properties computed with `Apply_prosail_inversion` are then 
saved as individual raster files. 
For each biophysical variable, the mean value and standard deviation defined by 
the input variable `nbModels` from `modelSVR` (default = 20 models) is saved.
