---
title: "2. Forward mode"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
  html_vignette:
    number_sections: true
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{2. Forward mode}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

## Run `prosail` in direct mode

### Input variables
The function `prosail` runs PROSAIL for individual canopy reflectances and 
expects the following input variables.

* `SAILversion`: the version of SAIL to be used. The default version is __4SAIL__. 
__4SAIL2__ can also be called from the same function with additional parameters. 
Both versions include the following input parameters:

* `spec_sensor`: dataframe including the refractive index and specific absorption 
coefficients, defined for a given spectral range (max range: 400 nm - 2500 nm). 
Simulation and inversion on different spectral domains can be performed by 
adapting the information in `SpecPROSPECT`, which is loaded by default when 
calling the package `prospect`.

* The biochemical and biophysical input variables of PROSPECT, including `N`, 
`chl`, `car`, `ant`, `brown`, `ewt`, `lma`, `prot`, `cbc`, `alpha`. 
If no value, their default value is set.

    * Note that a unique value (or no value) is expected for each of the leaf 
    constituents when running __4SAIL__. 
    __4SAIL2__ allows introduction of two distinct sets of leaf optical properties. 
    In the original version, one type is dedicated to 'Green leaves', and the 
    other type is dedicated to 'brown leaves'. 
    User can decide either to simulate both types of leaf optical properties 
    using PROSPECT, or to directly define these leaf optical properties with the 
    `brown_lop` input parameter, defined as follows

    * `brown_lop` is a parameter specific to  __4SAIL2__. 
    It should be a list including `Reflectance` and `Transmittance`, 
    with the spectral sampling corresponding to the sampling defined in `spec_sensor`
    
* `type_lidf`: two leaf inclination distribution functions are implemented here. 

    * Define `type_lidf=1` if you want to use the original LIDF introduced by 
    [Verhoef (1998)](https://edepot.wur.nl/210943).
        * if `type_lidf=1` then `lidf_a` controls the average leaf slope and 
        `lidf_b` controls the distribution's bimodality
        * This LIDF requires |lidf_a| + |lidf_b| < 1
| LIDF type    |   lidf_a   |   lidf_b   
| :----------: |:---------:|:---------:|
| Planophile   |     1     |     0     |
| Erectophile  |    -1     |     0     |
| Plagiophile  |     0     |    -1     |
| Extremophile |     0     |     1     |
| Spherical    |   -0.35   |   -0.15   |
| Uniform      |     0     |     0     |
    * Define `type_lidf=2` if you want to use the 
    [ellipsoidal distribution function](https://www.sciencedirect.com/science/article/pii/0168192386900109) 
    characterized by the average leaf inclination angle in degree
        * if `type_lidf=2` then `lidf_a` controls the average leaf angle

* `lai`: leaf area index

* `hotspot`: foliage hot spot parameter. It corresponds to the ratio of the 
correlation length of leaf projections in the horizontal plane and the canopy height [@doi:10.1016/j.rse.2006.12.013; @kuusk1985]

* `tts`: sun zenith angle (0 degrees = sun at the vertical of surface)

* `tto`: observer zenith angle (0 degrees = nadir observation)

* `psi`: azimuth between sun and observer (in degrees)

* `rsoil`: soil reflectance

Additionally, __4SAIL2__ also includes the following parameters

* `fraction_brown`: the fraction of LAI corresponding to brown leaf area, 
between 0 and 1

* `diss`: Layer dissociation factor

* `cv`: vertical crown cover percentage (= % ground area covered with crowns as 
seen from nadir direction)

* `zeta`: Tree shape factor (= ratio of crown diameter to crown height)

* `brown_lop`: optional input parameter. If user defined brown vegetation 
such as experimental measurements is available. If PROSPECT is used to define 
both green and brown leaves, leave `brown_lop` undefined, or set to `NULL`

### Output variables
`prosail` returns a list containing : 

* `rddt`: bi-hemispherical reflectance factor
* `rsdt`: directional-hemispherical reflectance factor for solar incident flux
*	`rdot`: hemispherical-directional reflectance factor in viewing direction
* `rsot`: bi-directional reflectance factor
* `fcover`: Fraction of green Vegetation Cover (= 1 - beam transmittance in the target-view path)
* `abs_dir`: canopy absorptance for direct solar incident flux
* `abs_hem`: canopy absorptance for hemispherical diffuse incident flux
* `rsdstar`: contribution of direct solar incident flux to albedo
* `rddstar`: contribution of hemispherical diffuse incident flux to albedo

```{r prosail direct mode}
library(prosail)
# define input variables for PROSPECT. 
# refer to prospect tutorial for default values
input_prospect <- data.frame('chl' = 40, 'car' = 8, 'ant' = 0.0, 
                             'ewt' = 0.01, 'lma' = 0.009, 'n_struct' = 1.5)
# define input variables for SAIL. 
lai <- 5      # LAI
hotspot <- 0.1      # Hot spot parameter
type_lidf <- 2 # leaf inclination distribution function 
lidf_a <- 30   # mean leaf angle
tts <- 30     # geometry of acquisition: sun zenith angle
tto <- 10     # geometry of acquisition: observer zenith angle
psi <- 90     # geometry of acquisition: sun-observer azimuth 
rsoil <- spec_soil_ossl$soil_01 # soil reflectance 
# - spec_soil: min_refl and max_refl
# - spec_soil_ossl: selection of 47 reflectance spectral representative of OSSL

# run prosail with 4SAIL
refl_prosail <- prosail(input_prospect = input_prospect, 
                      type_lidf = type_lidf, lidf_a = lidf_a, lai = lai,
                      hotspot = hotspot, tts = tts, tto = tto, psi = psi, rsoil = rsoil)

# run prosail with 4SAIL2
# define leaf chemical constituents corresponding green and brown leaves
input_prospect <- data.frame('chl' = c(40,5), 'car' = c(8,5), 'ant' = c(0,1), 
                             'ewt' = c(0.01,0.005), 'lma' = c(0.009,0.008), 
                             'brown' = c(0.0,0.5), 'n_struct' = c(1.5,2))
# define additional 4SAIL2 parameters
fraction_brown <- 0.5
diss <- 0.5
Cv <- 1
Zeta <- 1

# run prosail with 4SAIL2
refl_prosail2 <- prosail(SAILversion = '4SAIL2',
                       input_prospect = input_prospect,
                       type_lidf = type_lidf, lidf_a = lidf_a, lai = lai,
                       hotspot = hotspot, tts = tts, tto = tto, psi = psi, rsoil = rsoil,
                       fraction_brown = fraction_brown, diss = diss, 
                       cv = cv, zeta = zeta)
```


## Compute simplified bidirectional reflectance factor (BRF) under the assumption of clear conditions conditions

The function `compute_brf` computes the bi-directional reflectance factor in the 
direction of the observer, by combining both hemispherical-directional and 
bi-directional reflectance factors.
The direct and diffuse light are taken into account as described by 
[Francois et al. (2002)](https://doi.org/10.1051/agro:2002033), by computing the 
share of diffuse flux in global radiation `skyl` based on the equation proposed 
by [Spitters et al., 1986](https://www.sciencedirect.com/science/article/pii/0168192386900602).

```{r compute BRF}
# refl_prosail is the variable obtained when running prosail as in the previous 
# illustration.
# spec_atm corresponds to the direct and diffuse radiation solar spectra
BRF_4SAIL <- compute_brf(rdot = refl_prosail$rdot,
                         rsot = refl_prosail$rsot,
                         tts = tts,
                         spec_atm_sensor = spec_atm)
BRF_4SAIL2 <- compute_brf(rdot = refl_prosail2$rdot,
                          rsot = refl_prosail2$rsot,
                          tts = tts,
                          spec_atm_sensor = spec_atm)

```

## Compute fraction of absorbed photosyntehtically active radiation (fAPAR)

The function `compute_fapar` computes the fAPAR in the direction of the observer, 
by combining both canopy absorptance for direct solar incident flux and canopy 
absorptance for hemispherical diffuse incident flux. 

```{r compute fAPAR}
fapar_4sail <- compute_fapar(abs_dir = refl_prosail$abs_dir,
                             abs_hem = refl_prosail$abs_hem,
                             tts = tts,
                             spec_atm_sensor = spec_atm)
```

## Compute albedo

The function `compute_albedo` computes the albedo by combining the contribution 
of direct solar incident flux (rsdstar) and the contribution of hemispherical 
diffuse incident flux (rddstar). 

```{r compute albedo}
albedo_4SAIL <- compute_albedo(rsdstar = refl_prosail$rsdstar,
                               rddstar = refl_prosail$rddstar,
                               tts = tts,
                               spec_atm_sensor = spec_atm)
```

## Simulating sensor BRF

The simulation of sensor BRF requires the Spectral Response Function (SRF) of 
the sensor.

`prosail` includes the SRF corresponding to multiple sensors, including: 

* [Sentinel-2A and Sentinel-2B satellites](https://earth.esa.int/documents/247904/685211/S2-SRF_COPE-GSEG-EOPG-TN-15-0007_3.0.xlsx)

* [Landsat-7](https://landsat.usgs.gov/landsat/spectral_viewer/bands/L7_ETM_RSR.xlsx) and Landsat-8

* [MODIS](https://nwpsaf.eu/downloads/rtcoef_rttov12/ir_srf/rtcoef_eos_1_modis-shifted_srf.html)

* [Venus](https://labo.obs-mip.fr/multitemp/ven%c2%b5s-characteristics/)

* SPOT-6 and SPOT-7

```{r SRF 1}
# get the spectral response for Sentinel-2A
sensor_name <- 'Sentinel_2A'
# if interested in a different satellite, please use srf_path to 
# locate the SRF file expected to be named 'sensor_name_spectral_response.csv' 
# (separator = tabulations)
SRF_S2 <- get_radiometry(sensor_name,srf_path = NULL)
```

Other satellites can be simulated as long as the SRF is provided as CSV file, 
following the same 
[template](https://gitlab.com/jbferet/prosail/-/blob/master/data-raw/Sentinel_2_Spectral_Response.csv) 
as for the SRF of Sentinel-2 satellites. 

SRF can also be directly computed for any sensor under the hypothesis of 
Gaussian response for each spectral band, and information about the central 
wavelength and fwhm corresponding to each spectral band. 
The following example produces SRF corresponding to a hyperspectral sensor 
covering the spectral domain from 400 nm to 2400 nm with a fwhm of 10 nm for 
each channel.

```{r SRF 2}
# WL is a vector giving central wavelength for each band in nanometer
WL <- seq(400, 2400, by = 10)
# fwhm is a vector giving fwhm for each band in nanometer
fwhm <- 10+ 0*seq(400, 2400, by = 10)
# compute corresponding SRF
SRF_SH <- compute_srf(WL,fwhm)
```

Then, reflectance can be converted from 1-nm spectral sampling (default 
sampling) to sensor sampling. 
The following example uses SRF corresponding to Sentinel-2 sensor.

```{r Conversion}
# apply sensor characteristics to simulated reflectance
wvl <- spec_prospect_fullrange$lambda
BRF_4SAIL_S2 <- apply_sensor_characteristics(wvl = wvl,
                                           srf = SRF_S2,
                                           input_refl_table = BRF_4SAIL)
```
