---
title: "Tutorial for prospect"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```


This tutorial briefly describes the different functionalities of the R package __prospect__ based on the leaf physical model.
PROSPECT aims at simulating leaf optical properties (directional-hemispherical reflectance and transmittance) in the optical domain from 400 nm to 2500 nm based on their biophysical properties, including a limited number of biochemical constituents and a unique structure parameter, N. 

The current version of the model implemented in __prospect__ is PROSPECT-PRO, which includes the following biochemical constituents:

* Chlorophyll a + b
* Carotenoids
* Anthocyanins
* Prown pigments
* Equivalent water thickness (EWT)
* Leaf mass per area (LMA) 
* proteins
* carbon-based consituents (CBC, constituents of dry matter other than proteins)

The specific absoprtion coefficients corresponding to these constituents are recorded in the variable `SpecPROSPECT` available when loading the package __prospect__. 

# Running __prospect__ in direct mode

## Input variables
The function `PROSPECT` runs PROSPECT for individual samples and expects the following input variables.

* `SpecPROSPECT`: dataframe including the refractive index and specific absorption ceofficients, defined for a given spectral range (max range: 400 nm - 2500 nm). simulation and inversion on different spectral domains can be peformed by adapting the information in `SpecPROSPECT`
* The biochemical and biophysical input variables of PROSPECT can then be introduced following two possible ways:

    * as individual values : these input variables include `N`, `CHL`, `CAR`, `ANT`, `BROWN`, `EWT`, `LMA`, `PROT`, `CBC`, `alpha`. If no value, their default value is set
    * `Input_PROSPECT`: a dataframe including the full list of PROSPECT input parameters defind earlier. 

## Output variables
`PROSPECT` returns a list containing directional-hemispherical reflectance and transmittance (`reflectance` and `transmittance`) corresponding to the input variables.


```{r prospect direct mode}
library(biodivMapR)
#############################################################
###         run PROSPECT using default parameters         ###
#############################################################
LRT   = PROSPECT(SpecPROSPECT)

#############################################################
###   run PROSPECT using user defind set of parameters    ###
#############################################################
LRT2   = PROSPECT(SpecPROSPECT,N = 1.4,CHL = 30,CAR = 10,EWT = 0.02,LMA = 0.01)
```


# Computing a Look-Up-Table with __prospect__

The function `PROSPECT_LUT` allows computation of a LUT dirctly based on a list of input parameters


```{r prospect LUT}
#############################################################
###         run PROSPECT in order to produce a LUT        ###
### if some parameters are not set, their default value   ###
###               will be added automatically             ###
#############################################################
CHL = 100*runif(1000)
CAR = 25*runif(1000)
ANT = 2*runif(1000)
EWT = 0.04*runif(1000)
LMA = 0.02*runif(1000)
N   = 1+2*runif(1000)
Input_PROSPECT  = data.frame('CHL'=CHL,'CAR'=CAR,'ANT'=ANT,'EWT'=EWT,'LMA'=LMA,'N'=N)
LUT             = PROSPECT_LUT(SpecPROSPECT,Input_PROSPECT)
```


# Running __prospect__ in inverse mode

`PROSPECT` can be inverted using iterative optimization, by calling the function `Invert_PROSPECT`.
By default, the merit function used for the inversion minimizes RMSE bewen simulated and measured leaf optical properties. 
However, users can define their own merit function with associated criterion to minimize by defining their own merit function and adding it as input variable, such as `MeritFunction = MyOwnMeritFunction`

## Input variables
The function `Invert_PROSPECT` requires either reflectance, or transmittance, or both.
Use can define which input variable of `PROSPECT` should be estimated during inversion, and which ones should be set. 
The list of input variables for inversion is :

* `SpecPROSPECT`: dataframe including the refractive index and specific absorption ceofficients, defined for a given spectral range (max range: 400 nm - 2500 nm). simulation and inversion on different spectral domains can be peformed by adapting the information in `SpecPROSPECT`
* `Refl`: numeric: individual leaf reflectance corresponding to the spectral domain defined in `SpecPROSPECT`. Set to NULL if inversion on transmittance only
* `Tran`: numeric: individual leaf transmittance corresponding to the spectral domain defined in `SpecPROSPECT`. Set to NULL if inversion on reflectance only
* `Parms2Estimate` Parms2Estimate list. Parameters to estimate. Set to 'ALL' by default. the leaf structure parameter N is systematically estimated, unless defined in `ParmSet`
* `ParmSet` list of PROSPECT parameters which should be set to a given value
* `PROSPECT_version` character. corresponds to the PROSPECT version. should be one of the following versions: '5', '5B', 'D', 'DB', 'PRO', 'PROB'. Use the vesion ending with 'B' if you want to estimate brown pigments
* `MeritFunction` character. name of the function to be used as merit function with given criterion to minimize (default = RMSE)


## Output variables
`Invert_PROSPECT` returns a list containing estimated values of PROSPECT input parameters


